---
title: "Untitled"
author: "Andrea Termine"
date: '2022-07-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r config}
# plant a seed
set.seed(12345)

# Path to config
config = yaml::read_yaml(file = "../config/config r.yaml", eval.expr=TRUE)

# Paths
data.path       = file.path(config$exp$parent, config$file_05$datapath)
rna.meta.path   = file.path(config$exp$parent, config$file_05$rna_meta_dir)
lcf.out         = config$file_05$lcf.out
udf.path        = file.path(config$exp$parent, config$file_05$udf_dir)
vst.out         = config$file_05$vst.out
vst.ff.out      = config$file_05$vst.ff.out
vst.ff.stats.out= config$file_05$vst.ff.stats
```

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
# Import custom functions
source(file.path(udf.path, config$file_05$udf_name))

## Required Packages
pkg = c("tidyverse", "yaml", "fst",
        "biocManager","DESeq2", "M3C")

## Now load or install & load all // Cran version
load_packages(pkgs = pkg)
# cat("Libraries Imported")
rm(pkg,p)
```

# Aim

01 - Apply VST
02 - Apply center and scaling


# A note about the import

We tried to import the data using rds and parLapply or for each solutions. Since we had only two files they used only two cores and it was really slow. You could split in multiple small files or use the fst solution. Sources are listed below.

- [fst site](https://www.fstpackage.org/)
- [benchmark](https://waterdata.usgs.gov/blog/formats/)

```{r import}
# import metadata
rna_seq_metadata = readRDS(file.path(rna.meta.path, config$file_05$rna_meta_name))



# LCF counts import ----

## Prepare the import
count.paths = list.files(path       = lcf.out,
                         pattern    = ".fst",
                         full.names = T)

# storage object
count.list = vector(mode = "list", length = length(count.paths))
names(count.list) = list.files(path       = lcf.out,
                               pattern    = ".fst",
                               full.names = F) %>% word(sep = " ")

## Perform the import
pb = txtProgressBar(min = 0, max = length(count.list), initial = 0) 

for (i in 1:length(count.list)) {
  count.list[[i]] = read_fst(path = count.paths[i])
  
  setTxtProgressBar(pb = pb, value = i)
}
close(pb)
```

# VST Transformation

01 - Recreate the reproducible example from DESeq2 [V]
02 - Prepare counts for vst (wide format)          [V]
03 - Apply vst to our count matrices               [V]
04 - Apply feature filter after vst                [V]
05 - Create a vst_counts version with center/scale [V]
06 - Export vst and vst/centerscaled counts        [V]

You could create a dds object right now, but at this step I prefer to stick with the matrix/dataframe r objects. The variance stabilizing transformation is the main function that is usually called using vst. There are small differences that make the first function better if you will use it in Machine Learning pipeline. This kind of transformation is better than rlog for ML. You can find some sources below:

-[vst vs variance stabilizing transformation](https://www.biostars.org/p/459013/)
-[about data transformation and DESeq2](https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#data-transformations-and-visualization)

**Note on VST
There are three types of fits for the vst transformation. We will fit every estimator and then we will see if data clusterize and compare their silhouette.

```{r vst transformation Example}
# 01 ----
dds <- makeExampleDESeqDataSet(m=6)
vsd <- varianceStabilizingTransformation(dds)
dists <- dist(t(assay(vsd)))
plot(hclust(dists))

# inspect how a dds matrix should be built
assay(dds)
```

```{r vst transformation function with test}
# 02 Creaating a function for the transformations ----
test_genes = unique(count.list$hc_lcf_counts$Geneid)[1:50]
test       = count.list$hc_lcf_counts %>%  dplyr::filter(Geneid %in% test_genes) 

# Make a function
vst_transformer <- function(data, gene, id, counts, group_name, fit_types) {
  
  vars = c(id,gene,counts)
  
  print("Pivoting from long format")
  df = data %>%
    dplyr::select(all_of(vars)) %>%
    pivot_wider(names_from = vars[1], values_from = vars[3]) %>%
    column_to_rownames(var = vars[2]) %>%
    as.matrix()
  
  print("Pivoting from long format completed: Initializing storage objects")
  storage = vector(mode = "list", length = length(fit_types))
  labels  = paste(paste(group_name, paste("vst", fit_types, sep = "_"), sep = "_"), "counts", sep = "_")
  names(storage) = labels
  
  print("Storage object initialized: Starting the vst loop")
  for (i in 1:length(fit_types)) {
    print(paste("Starting VST type:", i))
    storage[[i]] = varianceStabilizingTransformation(object  = df,
                                                     blind   = TRUE,
                                                     fitType = fit_types[i]) %>%
      as.data.frame.matrix() %>% 
      rownames_to_column(var = gene)
  }

  return(storage)
}

# 2.45 sec elapsed for 50 genes and 188 subjects
tictoc::tic()
test_list = vst_transformer(test, gene = "Geneid", id = "PATNO", counts = "Count", group_name = "hc_lcf", fit_types = config$file_05$fit_types)
tictoc::toc()

# ---------------------------END oF TESTING-------------------------------------------#

# 03 Fit the transformer ----
vst.count.list = vector(mode = "list", length = length(count.list))

for (i in 1:length(count.list)) {
  print(paste("Starting: ", i))
  vst.count.list[[i]] =  vst_transformer(data       = count.list[[i]], 
                                         gene       = "Geneid",
                                         id         = "PATNO",
                                         counts     = "Count",
                                         group_name = config$file_05$group_names[i], 
                                         fit_types  = config$file_05$fit_types)
}

names(vst.count.list) = config$file_05$vst.count.list.names

# 03 Exports vst ----
for (i in 1:length(vst.count.list)) {
  for (ii in 1:length(vst.count.list[[i]])) {
    print(paste("Exporting:", names(vst.count.list[[i]])[[ii]]))
    
    write.fst(x        = vst.count.list[[i]][[ii]],
              path     = file.path(vst.out, paste0(names(vst.count.list[[i]])[[ii]],".fst")),
              compress = 100)
  }
}
```

```{r vst ff}
# 04 Create a vst + feature filter (ff) by MAD transformation ----
vst.ff.count.list = vst.count.list
vst.ff.stats.list = vst.count.list

for (i in 1:length(vst.count.list)) {
  for (ii in 1:length(vst.count.list[[i]])) {
    print(paste(i, ii))
    value = featurefilter(mydata     = vst.count.list[[i]][[ii]] %>% column_to_rownames("Geneid"),
                          percentile = 10,
                          method     = config$file_05$featurefilter,
                          topN       = 20)
    
    vst.ff.count.list[[i]][[ii]] = value[[1]] %>% rownames_to_column("Geneid") %>% relocate(Geneid, .before = everything())
    vst.ff.stats.list[[i]][[ii]] = value[[2]] %>%
      rownames_to_column("Geneid") %>% 
      relocate(Geneid, .before = everything()) %>%
      dplyr::select(-feature)
    
      names(vst.ff.count.list[[i]])[[ii]] = names(vst.ff.count.list[[i]])[[ii]] %>% 
      str_replace_all(pattern = "_counts", replacement = "_ff_counts")
  }
}

# 04 Exports vst + ff ----
for (i in 1:length(vst.ff.count.list)) {
  for (ii in 1:length(vst.ff.count.list[[i]])) {
    print(paste("Exporting:", names(vst.ff.count.list[[i]])[[ii]]))
    
    write.fst(x        = vst.ff.count.list[[i]][[ii]],
              path     = file.path(vst.ff.out, paste0(names(vst.ff.count.list[[i]])[[ii]],".fst")),
              compress = 100)
  }
}

saveRDS(vst.ff.stats.list, file = file.path(vst.ff.stats.out, "Feature Filtering stats.rds"))
```


```{r vst center scale}
# 05 Create a vst + center and scaling transformation ----
vst.centerscale.list = vst.count.list

for (i in 1:length(vst.centerscale.list)) {
  for (ii in 1:length(vst.centerscale.list[[i]])) {
    print(names(vst.centerscale.list[[i]])[ii])
    
    vst.centerscale.list[[i]][[ii]] = vst.centerscale.list[[i]][[ii]] %>%
      pivot_longer(cols       = !Geneid,
                   names_to   = "PATNO",
                   values_to  = "Count") %>% 
      pivot_wider(names_from  = "Geneid",
                  values_from = "Count") %>% 
      column_to_rownames("PATNO") %>%
      as.matrix() %>%
      scale(center = T, scale = T) %>% 
      as.data.frame.matrix() %>% 
      rownames_to_column("PATNO")
    
    names(vst.centerscale.list[[i]])[[ii]] = names(vst.centerscale.list[[i]])[[ii]] %>% 
      str_replace_all(pattern = "_counts", replacement = "_centerscale_counts")
  }
}

# 11 Exports vst + center/scaling ----
for (i in 1:length(vst.centerscale.list)) {
  for (ii in 1:length(vst.centerscale.list[[i]])) {
    print(paste("Exporting:", names(vst.centerscale.list[[i]])[[ii]]))
    
    write.fst(x        = vst.centerscale.list[[i]][[ii]],
              path     = file.path(vst.out, paste0(names(vst.centerscale.list[[i]])[[ii]],".fst")),
              compress = 100)
  }
}
```

```{r vst ff cs}
# 06 Create a vst + ff + center and scaling transformation ----
vst.ff.centerscale.list = vst.ff.count.list

for (i in 1:length(vst.ff.centerscale.list)) {
  for (ii in 1:length(vst.ff.centerscale.list[[i]])) {
    print(names(vst.ff.centerscale.list[[i]])[ii])
    
    vst.ff.centerscale.list[[i]][[ii]] = vst.ff.centerscale.list[[i]][[ii]] %>%
      pivot_longer(cols       = !Geneid,
                   names_to   = "PATNO",
                   values_to  = "Count") %>% 
      pivot_wider(names_from  = "Geneid",
                  values_from = "Count") %>% 
      column_to_rownames("PATNO") %>%
      as.matrix() %>%
      scale(center = T, scale = T) %>% 
      as.data.frame.matrix() %>% 
      rownames_to_column("PATNO")
    
    names(vst.ff.centerscale.list[[i]])[[ii]] = names(vst.ff.centerscale.list[[i]])[[ii]] %>% 
      str_replace_all(pattern = "_counts", replacement = "_centerscale_counts")
  }
}

# 08 Exports vst + ff + center/scaling ----
for (i in 1:length(vst.ff.centerscale.list)) {
  for (ii in 1:length(vst.ff.centerscale.list[[i]])) {
    print(paste("Exporting:", names(vst.ff.centerscale.list[[i]])[[ii]]))
    
    write.fst(x        = vst.ff.centerscale.list[[i]][[ii]],
              path     = file.path(vst.ff.out, paste0(names(vst.ff.centerscale.list[[i]])[[ii]],".fst")),
              compress = 100)
  }
}
```